### 2015-02-01 dot/sample.tex

\documentclass{standalone}
\usepackage[x11names, rgb]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{snakes,arrows,shapes}
\usepackage{amsmath}
%
%

%

%

\begin{document}
\pagestyle{empty}
%
%
%

% Start of code
% \begin{tikzpicture}[anchor=mid,>=latex',line join=bevel,]
\begin{tikzpicture}[>=latex',line join=bevel,]
  \pgfsetlinewidth{1bp}
%%
\pgfsetcolor{black}
  % Edge: N0 -- N1
  \draw [] (35.352bp,28.256bp) .. controls (55.506bp,30.438bp) and (88.555bp,30.437bp)  .. (108.69bp,28.252bp);
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (70.102bp,38.783bp) node {$l_1: 3Mbps$};
  % Edge: N0 -- N1
  \draw [] (35.352bp,18.027bp) .. controls (55.506bp,15.845bp) and (88.555bp,15.846bp)  .. (108.69bp,18.031bp);
  \draw (70.102bp,7.5bp) node {$l_2: 4Mbps$};
  % Node: N0
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (18.0bp,23.0bp) ellipse (18.0bp and 18.0bp);
  \draw (18.0bp,23.141bp) node {$a$};
\end{scope}
  % Node: N1
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (126.0bp,23.0bp) ellipse (18.0bp and 18.0bp);
  \draw (126.0bp,23.141bp) node {$b$};
\end{scope}
%
\end{tikzpicture}
% End of code

%
\end{document}
%